---
title: "README"
author: ""
date: "05/21/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Architecture of the folders

## code

This folder contains the codes for reproducing the analysis in Sections 5, 6 and the supplementary material of the paper (simulation studies and real data studies).

* **simul.R**: All results of simulation in the main manuscript (Section 5) can be reproduced by this file.
* **simul_supp.R**: All results of simulation in the supplementary materials (Section S9) can be reproduced by this file.
* **simul_CV.R**: Tuning parameters for simulation in the main manuscript (Section 5) can be selected by this file.
* **data_analysis.R**: All results of real data analysis in the main manuscript (Section 6) and the supplementary materials (Section S8) can be reproduced by this file.

There is also a sub-folder named "functions" which contains the codes for functions to implement our proposed methods and others.

* **spcovreg.R**: functions to implement our proposed method.
* **spcovreg_basic.R**: some simple functions such as the soft thresholding operator which have been used in spcovreg.R, simul.R, simul_supp.R, simul_CV.R and data_analysis.R.
* **covreg_em.R**: functions to implement the covariance regression by Hoff and Niu (2012).

## data

This folder contains three sub-folders for all the utilized data and output from both simulation and real data analyses.

* main: All data and output of simulation (Section 5) and real data analysis (Section 6) in the main manuscript are included in this folder. There are three types of RData files.
  * **CVresult_n200p50q30_MA_set1.RData**: This is the pregenerated list of tuning parameters ($\alpha$ and $\lambda$) selected by running the code "simul_CV.R". This file is loaded when running the code "simul.R" (see the code). Other files for different settings (e.g. $n=500$ and $q=100$) can be obtained by running "simul_CV.R" after modifying the code.
  * **fit_n200p50q30_MA_set1.RData**: This is the pregenerated set of results by running the code "simul.R". This file is loaded in Section 2 of this README file to produce numbers and figures. Other files for different settings (e.g. $n=500$ and $q=100$) can be obtained by running "simul.R" after modifying the code.
  * **gbmdata_run.RData**: This is the pregenerated set of results by running the code "data_analysis.R". This file is loaded in Section 2 of this README file to produce figures.
* supp: All the utilized data and output of simulation in the supplementary materials (Section S9) are included in this folder. There is one RData file in this folder.
  * **supp_q1.RData**: This is the pregenerated set of results by running the code "simul_supp.R". This file is loaded in Section 2 of this README file to produce numbers and figures.
* GBMdata: Raw data files for 178 patients with glioblastoma multiforme (GBM) from the REMBRANDT study (GSE108476). There are 6 raw files in this folder. The code for preprocessing these data can be found in "data_analysis.R".
  * **clinical.data.csv**: This data include age and gender data for the patients with GBM or other diseases. We considered only the GBM patients.
  * **gene names.csv**: This data contain the name of the genes considered.
  * **expData_73.RData**: This data contain the gene expression for the 73 genes considered.
  * **final_id.RData**: data for patient ID.
  * **SNPname.RData**: list of names for the 118 SNPs considered.
  * **SNPdata4.RData**: data for the 118 SNPs considered.

# 2. Reproducing the results in the manuscript and the supplementary materials
```{r}
################# set working directory #################
# please specify the path
path="/Users/rakheon_kim/Desktop/Research/SparseCovReg/"
```

## Numbers for tables: MA(1) model under Setting 1 with n=200, p=50, q=30
Numbers for other settings can be obtained by loading the R data for each setting.
```{r}
load(file=paste0(path,"data/main/fit_n200p50q30_MA_set1.RData"))
# Table 1
mean(S_diff); mean(ST_diff); mean(covreg_diff); mean(lse_diff); mean(covlpd_diff)
sd(S_diff); sd(ST_diff); sd(covreg_diff); sd(lse_diff); sd(covlpd_diff)
# Table 2
mean(sqrt(covl_err_est)); sd(sqrt(covl_err_est)); mean(lasso_TPR); mean(lasso_FPR)
# Table 3
colMeans(tb_COV)
```


## Figure 1 (Section 5)
```{r}
load(file=paste0(path,"data/main/fit_n500p50q30_MA_set1.RData"))
# Code for creating the picture can be found in simul.R
```

```{r, echo=FALSE}
par(mfrow=c(1,4), mar = c(2, 1, 2, 1))
plot(Sigma_12_mat[100,], S_12_mat[100,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[99,], S_12_mat[99,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[96,], S_12_mat[96,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[1,], S_12_mat[1,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[21,], S_12_mat[21,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
axis(2, at = c(-0.5,0.0,0.5,1.0,1.5), labels = NA)
axis(1, at = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5), labels = NA)

plot(Sigma_12_mat[100,], covreg_12_mat[100,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[99,], covreg_12_mat[99,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[96,], covreg_12_mat[96,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[1,], covreg_12_mat[1,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[21,], covreg_12_mat[21,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
axis(2, at = c(-0.5,0.0,0.5,1.0,1.5), labels = NA)
axis(1, at = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5), labels = NA)

plot(Sigma_12_mat[100,], lse_12_mat[100,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[99,], lse_12_mat[99,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[96,], lse_12_mat[96,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[1,], lse_12_mat[1,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[21,], lse_12_mat[21,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
axis(2, at = c(-0.5,0.0,0.5,1.0,1.5), labels = NA)
axis(1, at = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5), labels = NA)

plot(Sigma_12_mat[100,], covl_12_mat[100,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[99,], covl_12_mat[99,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[96,], covl_12_mat[96,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[1,], covl_12_mat[1,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
points(Sigma_12_mat[21,], covl_12_mat[21,], cex=0.1, ylim = c(-0.5,1.5), xlab="", ylab="", yaxt="n", xaxt="n")
axis(2, at = c(-0.5,0.0,0.5,1.0,1.5), labels = NA)
axis(1, at = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5), labels = NA)
```


## Figure 2 (Section 5)
```{r}
load(file=paste0(path,"data/main/fit_n500p50q100_MA_set2.RData"))
# Code for creating the picture can be found in simul.R
```

```{r, echo=FALSE}
library(ggplot2)
gfg<-round(data.frame(x = 1:(p*(p-1)/2),
                      y = covl_betaU[2,],
                      w = Sigma_off[2,],
                      low = covl_betaU_025[2,],
                      up = covl_betaU_975[2,]), 4)
nzid=which(Sigma_off[2,]!=0)
zid = nzid +1; zid = zid[1:48]
gfg4 = gfg[c(nzid, zid),]; gfg4$x=1:97
ggplot(gfg4, aes(x, w, y)) + geom_point(shape=1) + geom_point(aes(x,y)) + geom_errorbar(aes(ymin = low, ymax = up)) + scale_y_continuous(breaks=c(-1.5,-1,-0.5,0,0.5,1,1.5)) + theme_bw() + theme(axis.title = element_blank(),axis.text = element_text(size=10),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank(),legend.position="none")
```

## Figure 3 (Section 6)
```{r}
load(file=paste0(path,"data/main/gbmdata_run.RData"))
library(reshape2)
# Code for creating the picture can be found in data_analysis.R
```

```{r, fig.width=3.3, fig.height=3.3, echo=FALSE}
par(mfrow=c(1,2), mar = c(2, 1, 2, 1))
idsel=c( 64:66,58,61,62,57,72, 59,56,55, 25:27,15,13,17,22,16,20,21,23,24,12,14,18,19,48,49,50,45,46,47, 51:53,54,8,9,28,29,30,31,1, 40,44,43,42,41,36,37,39, 35,38,34,33,32,3,67,68,69,11,10,2,71,6,7,4,5, 73,70,63,60)
# left panel
Rn= ST[idsel,idsel]
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
# right panel
Rn= covl_coef[idsel,idsel,1]
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
```

## Figure 4 (Section 6)
```{r}
# Code for creating the picture can be found in data_analysis.R
```

```{r, fig.width=3.3, fig.height=3.3, echo=FALSE}
# left panel
Rn= covl_coef[idsel,idsel,18]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
# right panel
Rn= covl_coef[idsel,idsel,3]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
```

## Figure 5 (Section 6)
```{r}
library(network)
# Code for creating the picture can be found in data_analysis.R
# Depending on the version of R, the layout of the network graph may look differ
```

```{r, fig.width=3.3, fig.height=3.3, echo=FALSE}
#par(mfrow=c(1,2))
# left panel
mat = covl_coef[,,18]*nz_coef[,,18]
colnames(mat) = colnames(dat); rownames(mat) = colnames(dat)
matcol = (mat<0)*4 + (mat>0)*2
nn = as.network(mat, directed = FALSE)#, vertex.attr = mat)
set.seed(35)  # for XZ[,18]
lp = list(niter=NULL, max.delta=NULL, area=NULL, cool.exp=NULL, repulse.rad=NULL,
          init=NULL, groups=NULL, rotation=NULL, layout.control=0.5, constraints=NULL,
          round = TRUE, digits = 10)
lc = network.layout.fruchtermanreingold(nn, layout.par=lp)
par(mar = c(1, 1, 1, 1))
plot.network(nn, label = network.vertex.names(nn), coord = lc, label.cex = 0.6, vertex.cex = 0.5, edge.col = matcol, edge.lwd = abs(mat)*10)
# right panel
mat = covl_coef[,,3]*nz_coef[,,3]
colnames(mat) = colnames(dat); rownames(mat) = colnames(dat)
matcol = (mat<0)*4 + (mat>0)*2
nn = as.network(mat, directed = FALSE)#, vertex.attr = mat)
set.seed(28)  # for XZ[,3]
lp = list(niter=NULL, max.delta=NULL, area=NULL, cool.exp=NULL, repulse.rad=NULL,
          init=NULL, groups=NULL, rotation=NULL, layout.control=0.5, constraints=NULL,
          round = TRUE, digits = 10)
lc = network.layout.fruchtermanreingold(nn, layout.par=lp)
par(mar = c(1, 1, 1, 1))
plot.network(nn, label = network.vertex.names(nn), coord = lc, label.cex = 0.6, vertex.cex = 0.5, edge.col = matcol, edge.lwd = abs(mat)*10)
```


## Figure S1 (Section S8)
```{r}
# Code for creating the picture can be found in data_analysis.R
```

```{r, fig.width=3.3, fig.height=3.3, echo=FALSE}
# top left panel
Rn= covl_coef[idsel,idsel,94]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.4,0.4), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
# top right panel
Rn= covl_coef[idsel,idsel,33]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.4,0.4), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
# bottom left panel
Rn= covl_coef[idsel,idsel,28]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.1,0.1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
# bottom right panel
Rn= covl_coef[idsel,idsel,63]+diag(p)
rownames(Rn)=colnames(dat[,idsel]); colnames(Rn)=rev(colnames(dat[,idsel])); melted_R <- melt(Rn, na.rm = TRUE)
ggplot(data = melted_R, aes(Var1, rev(Var2), fill = value))+geom_tile(color = "white")+
  scale_fill_gradient2(low = "#0072B2", high = "#D55E00", mid = "white", midpoint = 0, limit = c(-0.1,0.1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal()+scale_x_discrete(position = "top") +
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=6, angle=90,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.background = element_blank(),
        legend.position="none") + coord_fixed()
```

## Figure S2 (Section S9)
```{r}
library(mvtnorm)
n=200; p=2; q=3; ii=1
set.seed(10000*ii)
z = matrix(runif(n*q), ncol=q)
z = cbind(z, z[,1]*z, z[,2]*z[,2:3], z[,3]*z[,3])
B0 = matrix(c(1,-1,1,1), ncol=2)
C0 = B0 %*% diag(c(1,1/3)) %*% t(B0) # diag(2) # B0 %*% diag(c(1,1/3)) %*% t(B0)
z1 = cbind(matrix(1,nrow=n,ncol=1), z[,1])
Sigma_indiv = array(0, dim=c(p,p,n))
dat = matrix(0, nrow = n, ncol = p)
for (i in 1:n){
  Sigma_indiv[1:2,1:2,i] = (C0/2) + (B0/2)%*%z1[i,]%*%t(z1[i,])%*%t(B0/2)
  dat[i,] = rmvnorm(1,mean=rep(0,p),sigma=Sigma_indiv[,,i])
}
par(mfrow=c(1,2), mar = c(2, 2, 2, 1))
plot(z[,1], Sigma_indiv[1,2,], col="white", pch=16, ylim=c(-6,2), xlab=NA, ylab=NA)
points(z[,1], (dat[,1]-mean(dat[,1]))*(dat[,2]-mean(dat[,2])), cex=0.5, pch=16)
plot(z[,2], Sigma_indiv[1,2,], col="white", pch=16, ylim=c(-6,2), xlab=NA, ylab=NA)
points(z[,2], (dat[,1]-mean(dat[,1]))*(dat[,2]-mean(dat[,2])), cex=0.5, pch=16)
```

## Numbers for tables in Section S9 (when q=1)
Numbers for other settings can be obtained by loading the R data for each setting.
```{r}
load(file=paste0(path,"data/supp/supp_q1.RData"))
# Table S1
mean(covreg_mcmc_diff); mean(covl_diff3); mean(covl_diff)
sd(covreg_mcmc_diff); sd(covl_diff3); sd(covl_diff)
```

## All of these can be obtained \url{https://github.com/rakheon/SparseCovReg}.